Parameter Inference\\with Bifurcation Diagrams

### 2021-06-11 docs/appendix.tex

\section{Properties of the Bifurcation Measure}
\label{appendix:measure}

>>>>

\section{Conditions for Bifurcations}
\label{appendix:conditions}

The conditions for bifurcations \eqref{eq:predictions} are expressed at a parametrised location $s\in\Reals$ on the equilibrium manifold. A local bifurcation \textit{not involving limit cycles} can be defined as a change of sign of the real part of an eigenvalue of the state-space Jacobian. In this section we show that the conditions
\begin{equation}
    \Det_{s=s^*} = 0
    \qquad
    \frac{d}{ds}\Det_{s=s^*} \neq 0
    \label{eq:conditions}
\end{equation}
are sufficient for a local bifurcation to occur at a specific location $s^*$ along the equilibrium manifold. Here we denote $\Det$ as the determinant of the state-space Jacobian. We begin by expressing the determinant is the product of eigenvalues
\begin{equation}
    \Det=\prod_{n=1}^N\lambda_n(s)
    \label{eq:determinant}
\end{equation}
Applying the product rule when differentiating yields
\begin{equation}
    \frac{d}{ds}\Det=
    \sum_{n=1}^N\frac{d\lambda_n}{ds}\prod_{n'\neq n}\lambda_{n'}(s)
\end{equation}
As $\Det_{s=s^*}=0$ then by \eqref{eq:determinant} one of the eigenvalues must vanish, say $\lambda^*(s)$ at  $s=s^*$ then
\begin{equation}
    \frac{d}{ds}\left|\frac{\partial\rates}{\partial u}\right|_{s=s^*}=
    \frac{d\lambda^*}{ds}\prod_{\lambda\neq\lambda^*}\lambda(s)\,\bigg|_{s=s^*}
\end{equation}
which implies the following conditions on the eigenvalues
\begin{align}
    \left.\lambda^*(s)\right|_{s=s^*}=0
    \qquad
    \left.\frac{d\lambda^*}{ds}\right|_{s=s^*}\neq 0
    \label{eq:sufficient-necessary}
    \\
    \lambda(s^*) \neq 0\quad\forall\lambda\neq\lambda*
    \qquad\quad
    \label{eq:sufficient}
\end{align}
Conditions \eqref{eq:sufficient-necessary} guarantee that eigenvalue $\lambda^*(s)$ crosses zero at $s^*$ with a finite slope and are therefore sufficient and necessary for a local non-degenerate bifurcation not involving limit cycles. Condition \eqref{eq:sufficient} restricts the bifurcation at $s^*$ to be such that only one eigenvalue may vanish with all others being finite. Therefore conditions \eqref{eq:conditions} are sufficient \textit{but not necessary} for local non-degenerate bifurcations at $s^*$ not involving limit cycles.

In the case where more than one eigenvalue vanishes at the bifurcation the second condition \eqref{eq:conditions} is broken. This is because the determinant scales like $(s-s^*)^k$ where $k$ is the number of eigenvalues crossing zero. When the second condition \eqref{eq:conditions} is broken the measure \eqref{eq:measure} becomes zero. This means that the unsupervised term does not encourage bifurcations which have more than one eigenvalue crossing zero.
